---
title: "Tidy exercises in class"
author: "Tianying Zhang"
date: "2/21/2018"
output: pdf_document
---

# Tidy Data


## Introduction

```{r message=FALSE}
library(tidyverse)
```

## Tidy Data

1. Using prose, describe how the variables and observations are organized in each of the sample tables.

In `table1` each row is a (country, year) with variables `cases` and `population`.
```{r}
table1
```

In `table2`, each row is country, year , variable ("cases", "population") combination, and there is a `count` variable with the numeric value of the variable.
```{r}
table2
```

In `table3`, each row is a (country, year) combination with the column `rate` having the rate of cases to population as a character string in the format `"cases/rate"`.
```{r}
table3
```

Table 4 is split into two tables, one table for each variable: `table4a` is the table for cases, while `table4b` is the table for population. Within each table, each row is a country, each column is a year, and the cells are the value of the variable for the table.
```{r}
table4a
```
```{r}
table4b
```

2. Compute the `rate` for `table2`, and `table4a` + `table4b`. You will need to perform four operations:

  * Extract the number of TB cases per country per year.
  * Extract the matching population per country per year.
  * Divide cases by population, and multiply by 10000.
  * Store back in the appropriate place.
  
Which representation is easiest to work with? Which is hardest? Why?

Without using the join functions introduced in Ch 12:
```{r}
tb2_cases <- filter(table2, type == "cases")[["count"]]
tb2_country <- filter(table2, type == "cases")[["country"]]
tb2_year <- filter(table2, type == "cases")[["year"]]
tb2_population <- filter(table2, type == "population")[["count"]]
table2_clean <- tibble(country = tb2_country,
       year = tb2_year,
       rate = tb2_cases / tb2_population)
table2_clean
```
Note, that this assumes that all observations are sorted so that each country, year will have the observation for cases followed by population.

```{r}
tibble(country = table4a[["country"]],
       `1999` = table4a[["1999"]] / table4b[["1999"]],
       `2000` = table4b[["2000"]] / table4b[["2000"]])
```
or 
```{r}
tibble(country = rep(table4a[["country"]], 2),
       year = rep(c(1999, 2000), each = nrow(table4a)),
       `rate` = c(table4a[["1999"]] / table4b[["1999"]],
                  table4b[["2000"]] / table4b[["2000"]]))
```


3. Recreate the plot showing change in cases over time using `table2` instead of `table1`. What do you need to do first?

First, I needed to filter the tibble to only include those rows that represented the "cases" variable.
```{r}
table2 %>% 
  filter(type == "cases") %>%
  ggplot(aes(year, count)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```


## Spreading and Gathering

This code is reproduced from the chapter because it is needed by the exercises:
```{r}
tidy4a <- table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
```


1. Why are `gather()` and `spread()` not perfectly symmetrical?
   Carefully consider the following example:

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```

The functions `spread` and `gather` are not perfectly symmetrical because column type information is not transferred between them. 
In the original table the column `year` was numeric, but after running `spread()` and `gather()` it is a character vector. 
This is because variable names are always converted to a character vector by `gather()`.

The `convert` argument tries to convert character vectors to the appropriate type. 
In the background this uses the `type.convert` function.
```{r}
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`, convert = TRUE)
```

2. Why does this code fail?

```{r error=TRUE}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
```

The code fails because the column names `1999` and `2000` are not standard and thus needs to be quoted.
The tidyverse functions will interpret `1999` and `2000` without quotes as looking for the 1999th and 2000th column of the data frame.
This will work:
```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```

3. Why does spreading this tibble fail? How could you add a new column to fix the problem?


```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
glimpse(people)
```

```{r error=TRUE}
spread(people, key, value)
```

Spreading the data frame fails because there are two rows with "age" for "Phillip Woods".
We would need to add another column with an indicator for the number observation it is,

```{r}
people <- tribble(
  ~name,             ~key,    ~value, ~obs,
  #-----------------|--------|------|------
  "Phillip Woods",   "age",       45, 1,
  "Phillip Woods",   "height",   186, 1,
  "Phillip Woods",   "age",       50, 2,
  "Jessica Cordero", "age",       37, 1,
  "Jessica Cordero", "height",   156, 1
)
spread(people, key, value)
```


4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

You need to gather it. The variables are:

- pregnant: logical ("yes", "no")
- female: logical
- count: integer

```{r}
gather(preg, sex, count, male, female) %>%
  mutate(pregnant = pregnant == "yes",
         female = sex == "female") %>%
  select(-sex)
```
Converting `pregnant` and `female` from character vectors to logical was not necessary to tidy it, but it makes it easier to work with.