---
title: "HW#4"
author: "Tianying Zhang"
date: "2/26/2018"
output: pdf_document
---
#12.6.1

```{r message=FALSE}
library("tidyverse")
```
## Case Study

```{r}
who1 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
glimpse(who1)
```

```{r}
who2 <- who1 %>%
 mutate(key = stringr::str_replace(key, "newrel", "new_rel"))

```

```{r}
who3 <- who2 %>%
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```

```{r}
who3 %>%
  count(new)
```

```{r}
who4 <- who3 %>%
  select(-new, -iso2, -iso3)
```

```{r}
who5 <- who4 %>%
  separate(sexage, c("sex", "age"), sep = 1)
who5
```


3. I claimed that `iso2` and `iso3` were redundant with country. Confirm this claim.

```{r}
select(who3, country, iso2, iso3) %>%
  distinct() %>%
  group_by(country) %>%
  filter(n() > 1)

```


4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualization of the data.

```{r}
who5 %>%
  group_by(country, year, sex) %>%
  filter(year > 1995) %>%
  summarise(cases = sum(cases)) %>%
  unite(country_sex, country, sex, remove = FALSE) %>%
  ggplot(aes(x = year, y = cases, group = country_sex, colour = sex)) +
  geom_line()

  
```

A small multiples plot faceting by country is difficult given the number of countries.
Focusing on those countries with the largest changes or absolute magnitudes after providing the context above is another option.

#10.5

# Tibbles

```{r message=FALSE}
library("tidyverse")
```


## Exercises
5. What does `tibble::enframe()` do? When might you use it?

It converts named vectors to a data frame with names and values
```{r eval=FALSE}
?tibble::enframe
```

```{r}
enframe(c(a = 1, b = 2, c = 3))
```

## table 4 -> table 6

```{r}

library(foreign)
library(stringr)
library(plyr)
library(reshape2)
source("xtable.r")

# Data from http://pewforum.org/Datasets/Dataset-Download.aspx

# Load data -----------------------------------------------------------------

pew <- read.spss("pew.sav")
pew <- as.data.frame(pew)


religion <- pew[c("q16", "reltrad", "income")]
religion$reltrad <- as.character(religion$reltrad)
religion$reltrad <- str_replace(religion$reltrad, " Churches", "")
religion$reltrad <- str_replace(religion$reltrad, " Protestant", " Prot")
religion$reltrad[religion$q16 == " Atheist (do not believe in God) "] <- "Atheist"
religion$reltrad[religion$q16 == " Agnostic (not sure if there is a God) "] <- "Agnostic"
religion$reltrad <- str_trim(religion$reltrad)
religion$reltrad <- str_replace_all(religion$reltrad, " \\(.*?\\)", "")

religion$income <- c("Less than $10,000" = "<$10k", 
                     "10 to under $20,000" = "$10-20k", 
                     "20 to under $30,000" = "$20-30k", 
                     "30 to under $40,000" = "$30-40k", 
                     "40 to under $50,000" = "$40-50k", 
                     "50 to under $75,000" = "$50-75k",
                     "75 to under $100,000" = "$75-100k", 
                     "100 to under $150,000" = "$100-150k", 
                     "$150,000 or more" = ">150k", 
                     "Don't know/Refused (VOL)" = "Don't know/refused")[religion$income]

religion$income <- factor(religion$income, levels = c("<$10k", "$10-20k", "$20-30k", "$30-40k", "$40-50k", "$50-75k", 
                                                      "$75-100k", "$100-150k", ">150k", "Don't know/refused"))

counts <- count(religion, c("reltrad", "income"))
names(counts)[1] <- "religion"

raw <- dcast(counts, religion ~ income)
raw
raw2 <- filter(raw, religion == "Agnostic")

raw2 %>% gather(key = "income", value = "freq", -religion)
```
## table 7 -> table 8
```{r}
options(stringsAsFactors = FALSE)
library(lubridate)
library(reshape2)
library(stringr)
library(plyr)
source("xtable.r")

raw <- read.csv("billboard.csv")
raw <- raw[, c("year", "artist.inverted", "track", "time", "date.entered", 
               "x1st.week", "x2nd.week", "x3rd.week", "x4th.week", "x5th.week", 
               "x6th.week", "x7th.week", "x8th.week", "x9th.week", "x10th.week", 
               "x11th.week", "x12th.week", "x13th.week", "x14th.week", "x15th.week", 
               "x16th.week", "x17th.week", "x18th.week", "x19th.week", "x20th.week", 
               "x21st.week", "x22nd.week", "x23rd.week", "x24th.week", "x25th.week", "x26th.week", "x27th.week", "x28th.week", "x29th.week", "x30th.week", "x31st.week", "x32nd.week", "x33rd.week", "x34th.week", "x35th.week", "x36th.week", "x37th.week", "x38th.week", "x39th.week", "x40th.week", "x41st.week", "x42nd.week", "x43rd.week", "x44th.week", "x45th.week", "x46th.week", "x47th.week", "x48th.week", "x49th.week", "x50th.week", "x51st.week", "x52nd.week", "x53rd.week", "x54th.week", "x55th.week", "x56th.week", "x57th.week", "x58th.week", "x59th.week", "x60th.week", "x61st.week", "x62nd.week", "x63rd.week", "x64th.week", "x65th.week", "x66th.week", "x67th.week", "x68th.week", "x69th.week", "x70th.week", "x71st.week", "x72nd.week", "x73rd.week", "x74th.week", "x75th.week", "x76th.week")]
names(raw)[2] <- "artist"

raw$artist <- iconv(raw$artist, "MAC", "ASCII//translit")
raw$track <- str_replace(raw$track, " \\(.*?\\)", "")
names(raw)[-(1:5)] <- str_c("wk", 1:76)
raw <- arrange(raw, year, artist, track)

long_name <- nchar(raw$track) > 20
raw$track[long_name] <- paste0(substr(raw$track[long_name], 0, 20), "...")
head(raw,10)
clean <- melt(raw, id = 1:5, na.rm = T)
clean$week <- as.integer(str_replace_all(clean$variable, "[^0-9]+", ""))
clean$variable <- NULL

clean$date.entered <- ymd(clean$date.entered)
clean$date <- clean$date.entered + weeks(clean$week - 1)
clean$date.entered <- NULL
clean <- rename(clean, c("value" = "rank"))
clean <- arrange(clean, year, artist, track, time, week)
clean <- clean[c("year", "artist", "time", "track", "date", "week", "rank")]

clean_out <- mutate(clean, 
                    date = as.character(date))
head(clean_out,10)
```
